The Driving Forces of API Evolution

### 2015-05-25 thesis/paper.tex
meth table
Small changes in meth

\begin{table}
       \centering
       \begin{tabular}[ht]{l|c}
              \toprule

@@

\begin{table}
       \centering
       \begin{tabular}[ht]{p{3in}}
       \toprule
              \begin{enumerate}
              \item \textbf{Continuing Change}  \newline
              E-Type systems must be continually adapted else they become progressively less satisfactory.                     
              \item \textbf{Increasing Complexity} \newline
              As an E-Type system evolves its complexity increases unless work is done to maintain or reduce it.                  
              \item \textbf{Self Regulation} \newline
              E-Type system evolution process is self regulating with distribution of product and process measures close to normal.                     
              \item \textbf{Conservation of Organisational Stability} \newline
              The average effective global activity rate in an evolving E-Type system is invariant over product lifetime.                     
              \item \textbf{Conservation of Familiarity} \newline
              As an E-type system evolves all associated with it, developers, sales personnel, users, for example, must maintain mastery of its content and behaviour to achieve satisfactory evolution. Excessive growth diminishes that mastery. Hence the average incremental growth remains invariant as the system evolves.                     
              \item \textbf{Continuing Growth} \newline
              The functional content of E-Type systems must be continually increased to maintain user satisfaction over their lifetime.                
              \item \textbf{Declining Quality} \newline
              The quality of E-Type systems will appear to be declining unless they are rigorously maintained and adapted to operational environment changes.                     
              \item \textbf{Feedback System} \newline
              E-Type evolution process constitute multi-level, multi-loop, multi-agent feedback systems and must be treated as such to achieve significant improvement over any reasonable base.

@@

API design is notoriously difficult, a myriad of design and performance decisions must be taken into consideration when creating APIs \cite{afonso2012evaluating, bloch2008effective, stylos2006comparing}. Examples of such design decisions include how to structure an object's constructor parameters or if the API should display errors at compilation or at runtime \cite{stylos2006comparing}. More trivial design problems, such as assigning names to API features or correctly naming user-defined types, can have a significant impact on the usability of an API \cite{shi2011empirical}. When facing such design challenges, the following four factors are important to consider: a) The API must be understandable through good documentation, b) the API must not be overly abstract, c) the API must be reusable and d) the API must be easy to learn \cite{shi2011empirical}. One of the most important qualities in an API is that the intent of the API must be clear to the user \cite{shi2011empirical, stylos2006comparing}. The design decisions reached during development of an API will affect the overall usability of the API. Measuring such an effect can be done by investigating the twelve cognitive dimensions that are impacted by interactions between the API and its users \cite{clarke2004measuring}.

@@

              \textbf{Phase 1: Generate Hypotheses}

              \begin{enumerate}
				\item Inspect 1.4
				\item Inspect 2.0
				\item Identify changes between versions
				\item Conduct interview
				\item Structure gathered data
              \end{enumerate} \\

              \textbf{Phase 2: Confirm Hypotheses}

              \begin{enumerate}
              \item Code changes
              \item Group changes
              \item Categorise groups 
              \item Identify and formulate trends
              \item Analyse based on types \cite{chapin2001types}
              \item Analyse based on laws \cite{lehman90sview}
              \end{enumerate} \\

              \bottomrule
       \end{tabular}
       \caption{Methodology Steps}
       \label{table:meth_steps}
\end{table}

@@

\subsection{Data Collection} \label{data_collection}  
%Intro to code inspection
The data were collected primarily through inspections of the source code and the analysis of accompanying documentation. The source code was comprised of the two API versions previously mentioned. Each version of the API had its own respective documentation, including example code, a library description and a basic development manual.

@@

%Purpose of the interview
After the initial source code inspection and preliminary analysis had been performed, an interview was conducted. The main purpose of the interview was to validate our findings from the code inspection with the API architect. The interview also served the purpose of providing additional insight into what drove the evolution of the API, by complementing what was found from the code inspection. The final reason for interviewing the API architect was to gather information about the API users and how their decisions, needs and requirements have affected the development of the API.

%Structure of the interview
The interview was conducted using a semi-structured approach \cite{robson2002real}. Structure to the interview was provided by organising it around the software evolution theories \cite{chapin2001types, lehman90sview}. Investigative and open-ended questions related to each type and law were asked to provide a basis for comparison between the answers and our initial analysis. The interviewee was encouraged to speak freely, even if it required a change in direction or topic. This was allowed in order to fulfil our exploratory goal of the study.

@@

%RQ1
The trends identified previously were then classified with help of the decision tree used for classifying the types of software evolution and software maintenance \cite{chapin2001types}. Each of the previously identified trends were given a main type of change and, if appropriate, a secondary type. Main types of change identify the primary motive for a trend, whereas the secondary types are additional motivations for a trend. For example, if an identified trend was the re-implementation of an old module that included new functionality and made use of a new interface, it would be classified as mainly enahnacive and secondarily groomative. In addition to this, we analysed the trends based on Lehman's Laws, to determine if they applied to the changes made to the API. This material, as well as the results of the interview, were then used as the basis for answering \textbf{RQ1}.

>>>>

API design is notoriously difficult, a myriad of design and performance decisions must be taken into consideration when creating APIs \cite{afonso2012evaluating, bloch2008effective, stylos2006comparing}. Examples of such design decisions include how to structure an object's constructor parameters or if the API should display errors at compilation or at runtime \cite{stylos2006comparing}. More trivial design problems, such as assigning names to API features or correctly naming user-defined types, can have a significant impact on the usability of an API \cite{shi2011empirical}. When facing such design challenges, the following four factors are important to consider: a) The API must be understandable through good documentation, b) the API must not be overly abstract, c) the API must be reusable and d) the API must be easy to learn \cite{shi2011empirical}. One of the most important qualities in an API is that the intent of the API must be clear to the user \cite{shi2011empirical, stylos2006comparing}. The design decisions reached during development of an API will affect the overall usability of the API. Measuring such an effect can be done by investigating the twelve cognitive dimensions that are impacted by interactions between the API and its users \cite{clarke2004measuring}. 
\begin{table}
       \centering
       \begin{tabular}[Ht*]{l|c}
              \toprule

@@

\begin{table}
       \centering
       \begin{tabular}[Ht*]{p{3in}}
       \toprule
              \begin{enumerate}
              \item \textbf{Continuing Change}  \newline
              E-Type systems must be continually adapted else they become progressively less satisfactory.                     
              \item \textbf{Increasing Complexity} \newline
              As an E-Type system evolves its complexity increases unless work is done to maintain or reduce it.                  
              \item \textbf{Self Regulation} \newline
              E-Type system evolution process is self regulating with distribution of product and process measures close to normal.                     
              \item \textbf{Conservation of Organisational Stability} \newline
              The average effective global activity rate in an evolving E-Type system is invariant over product lifetime.                     
              \item \textbf{Conservation of Familiarity} \newline
              As an E-type system evolves all associated with it, developers, sales personnel, users, for example, must maintain mastery of its content and behaviour to achieve satisfactory evolution. Excessive growth diminishes that mastery. Hence the average incremental growth remains invariant as the system evolves.                     
              \item \textbf{Continuing Growth} \newline
              The functional content of E-Type systems must be continually increased to maintain user satisfaction over their lifetime.                
              \item \textbf{Declining Quality} \newline
              The quality of E-Type systems will appear to be declining unless they are rigorously maintained and adapted to operational environment changes.                     
              \item \textbf{Feedback System} \newline
              E-Type evolution process constitute multi-level, multi-loop, multi-agent feedback systems and must be treated as such to achieve significant improvement over any reasonable base.

@@

@@

              \textbf{Phase 1: Data Collection}

              \begin{enumerate}
				\item Inspect v1.4
				\item Inspect v2.0
				\item Identify changes between versions
				\item Conduct interview
				\item Structure gathered data
				\item Formulate Hypotheses
              \end{enumerate} \\

              \textbf{Phase 2: Data Analysis}

              \begin{enumerate}
              \item Code changes
              \item Group changes
              \item Categorise groups 
              \item Identify and formulate trends
              \item Analyse based on types \cite{chapin2001types}
              \item Analyse based on laws \cite{lehman90sview}
              \item Validate hypotheses
              \end{enumerate} \\

              \bottomrule
       \end{tabular}
       \caption{Methodology Overview}
       \label{table:meth_steps}
\end{table}

@@

\subsection{Data Collection} \label{data_collection}  
%Intro to code inspection
The data were collected primarily through inspections of the source code and of accompanying documentation. The source code was comprised of the two API versions previously mentioned. Each version of the API had its own respective documentation, including example code, a library description and a basic development manual.

@@

%Purpose of the interview
After the initial source code and documentation inspection, an interview was conducted. The main purpose of the interview was to validate our findings from the code inspection with the API architect. The interview also served the purpose of providing additional insight into what drove the evolution of the API, by complementing what was found from the inspections. The final reason for interviewing the API architect was to gather information about the API users and how their decisions, needs and requirements have affected the development of the API.

%Structure of the interview
The interview was conducted using a semi-structured approach \cite{robson2002real}. Structure to the interview was provided by organising it around the software evolution theories \cite{chapin2001types, lehman90sview}. Investigative and open-ended questions related to each type and law were asked to provide a basis for comparison between the answers and our initial analysis. The interviewee was encouraged to speak freely, even if it required a change in direction or topic. This was encouraged in order to fulfil our exploratory goal of the study.

@@

%RQ1
The trends identified previously were then classified with help of the decision tree used for classifying the types of software evolution and software maintenance \cite{chapin2001types}. Each of the previously identified trends were given a main type of change and, if appropriate, a secondary type. Main types of change identify the primary motive for a trend, whereas the secondary types are additional motivations for a trend. For example, if an identified trend was the re-implementation of an old module that included new functionality and made use of a new interface, it would be classified as mainly enahnacive and secondarily groomative. In addition to this, we analysed the trends based on Lehman's Laws, to determine if they applied to the changes made to the API. This information, as well as the results of the interview, were then used as the basis for answering \textbf{RQ1}.
